---
title: "Operation on Vectors"
output: html_notebook
---

What is vectors?
A vector is a sequence of data elements of the same basic type. 

### A. How to create vector in R?
There are numerous ways to create an R vector.

#### Using c() Function
```{r}
vec <- c(1,2,3,4,5)
vec
```

#### Using assign() function
```{r}
assign("vec2",c(6,7,8,9,10)) 
vec2
```

#### Using : operator
```{r}
vec3 <- 1:20
vec3
```

### B. Basic Operation on Vector
Basic operation such as arithmetic operations of vectors are performed memberwise (member-by-member).
Suppose we have two vectors a and b.
```{r}
a = c(1, 3, 5, 7) 
b = c(1, 2, 4, 8)
```

#### Scalar Multiplication
```{r}
5 * a 
```

#### Vector addition
```{r}
a + b
```

#### Vector subtraction
```{r}
(3*a) - b
```

#### Vector multiplication
```{r}
3*a * b
```

### C. Norm and Distance of Vectors
#### Vector Norm
Length of a vector also called as *norm*.
```{r}
v <- c(2, 3, -1, 2)
v  
```
We can define our own norm function.
```{r}
norm_vec <- function(x) 
  sqrt(sum(x^2))
```

```{r}
norm_vec(v)   #Calculate norm of a vector
```

There is also a norm function in R.
```{r}
vec_norm <- norm(v, type = "2")    # Why type = '2'??
vec_norm  
```

#### Vectors Distance
Calculate the distance between two vector
```{r}
w <- c(-2, 3, 3, -1)
```

We can fefine our own distance function.
```{r}
euclidean <- function(a, b) sqrt(sum((a - b)^2))
```

```{r}
euclidean(v,w)
```

There is also the dist function in R.
```{r}
dist(rbind(v,w), method = "euclidean")
```

### D. Dot Product
The dot product or also known as the scalar product is an algebraic operation that takes two equal-length sequences of numbers and returns a single number.

Import the required library 
```{r}
library(geometry) 
```

#### Example 1: Taking two scalar values 
```{r}
a = 5 
b = 7 
print(dot(a, b, d = TRUE))
```

#### Example 2: Taking two simple vectors 
```{r}
a = c(1, 4) 
b = c(7, 4) 
print(dot(a, b))
```

#### Example 3: Taking two 2D array 
```{r}
vector1 = c(2, 1) 
vector2 = c(0, 3) 
a = array(c(vector1, vector2), dim = c(2, 2)) 
a

vector1 = c(4, 2) 
vector2 = c(9, 3) 
b = array(c(vector1, vector2), dim = c(2, 2)) 
b

print(dot(a, b))
```

### E. Cross Product
The cross product or also known as the vector product is a binary operation on two vectors in three-dimensional space and is denoted by the symbol ‘X‘. 

Given two linearly independent vectors a and b, the cross product, *a × b* is a vector that is perpendicular to both a and b and thus normal to the plane containing them.

Import the required library 
```{r}
library(pracma) 
```

#### Example 1: Taking input as vectors
```{r}
a = c(3, 5, 4) 
b = c(2, 7, 5) 
print(cross(a, b))
```

#### Example 2: Taking input as matrix
```{r}
a = matrix(  
  c(1, 2, 3, 4, 5, 6, 7, 8, 9),   
  nrow = 3,               
  ncol = 3,               
  byrow = TRUE           
)  
a

b = matrix(  
  c(5, 2, 1, 4, 6, 6, 3, 2, 9),   
  nrow = 3,               
  ncol = 3,               
  byrow = TRUE           
)  
b

print(cross(a, b))
```

#### Example 3: Taking input matrix 4x4
```{r}
c = matrix(  
  c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16),   
  nrow = 4,               
  ncol = 4,               
  byrow = TRUE           
)  
c

d = matrix(  
  c(5, 2, 1, 4, 6, 6, 3, 2, 9, 11, 13, 15, 17, 19, 21, 23),   
  nrow = 4,               
  ncol = 4,               
  byrow = TRUE           
)  
d

print(crossprod(c, d))
```

